Analyze the logical forms of the following statements:
\begin{enumerate}[label=(\alph*)]
    \item Mary will sell her house only if she can get a good price and find a nice apartment.
    \item Having both a good credit history and an adequate down payment is a necessary condition for getting a mortgage.
    \item John will kill himself, unless someone stops  him. (Hint: First try to rephrase this using the words \textit{if} and \textit{then} instead of \textit{unless}.)
    \item If $x$ is divisible by either 4 or 6, then it isn't prime.
\end{enumerate}

\textbf{Solution:}
\begin{enumerate}[label=(\alph*)]
    \item Let $S$ stand for the statement "\textbf{May will sell her house.}" Let $P$ stand for "\textbf{The house will sell for a good price.}" Finally, let $A$ stand for "\textbf{May can find a nice apartment.}" Then the logical form of the given statement can be expressed as $S \rightarrow (P \wedge A)$.
    \item Let $C$ stand for the statement "\textbf{Have a good credit history.}" Let $D$ stand for the statement "\textbf{Have an adequate down payment.}" Let $M$ stand for "\textbf{Having a mortgage.}" The logical form of the given statement is $M \rightarrow (C \wedge D)$.
    \item Let $K$ stand for the statement "\textbf{John will kill himself.}" Let $S$ stand for "\textbf{Someone will not stop John from killing himself.}" The logical form of the given statement is $\neg S \rightarrow K$.
    \item Let $A(x)$, $B(x)$, and $C(x)$ stand for the statements: "\textbf{\bm{$x$} is divisible by 4}", "\textbf{\bm{$x$} is divisible by 6}", and "\textbf{\bm{$x$} is a prime number}," correspondingly. The logical form of the given statement is $[A(x) \vee B(x)] \rightarrow \neg \: C(x)$.
\end{enumerate}
\pagebreak